---
title: "Swiss Fertility Dataset — Correlation Matrix"
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: true
    embed-resources: true
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(corrplot)
```

## Load the data

The `swiss` dataset (built into R) records socioeconomic indicators for 47 French-speaking cantons of Switzerland around 1888.

```{r}
#| label: load-data
data(swiss)
glimpse(swiss)
```

## Correlation matrix

```{r}
#| label: corr-plot
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Correlations among the six variables in the swiss dataset (Pearson)."
cor_mat <- cor(swiss)

corrplot(
  cor_mat,
  method   = "color",
  type     = "full",
  order    = "hclust",
  tl.col   = "black",
  tl.srt   = 45,
  tl.cex   = 1.1,
  addCoef.col = "black",
  number.cex  = 0.9,
  col = colorRampPalette(c("#BB4444", "#FFFFFF", "#4477AA"))(200),
  title = "Pearson correlations: swiss dataset",
  mar   = c(0, 0, 2, 0)
)
```

```{r}
#| label: corr-table
#| echo: false
knitr::kable(
  round(cor_mat, 2),
  align = "c",
  caption = "Correlation coefficients (rounded to 2 decimals)"
)
```

## Interpretation of each correlation

Each sentence below covers one of the 15 unique variable pairs (the diagonal is 1 by definition).

### Fertility

- **Fertility ~ Agriculture (r = 0.35):** Cantons with a larger share of the workforce in agriculture tend to have moderately higher fertility, consistent with more rural, traditional communities.
- **Fertility ~ Examination (r = −0.65):** Cantons that score higher on the army examination (a proxy for recruit quality/education) have substantially lower fertility.
- **Fertility ~ Education (r = −0.66):** Higher literacy/education rates are strongly associated with lower fertility, a classic demographic pattern.
- **Fertility ~ Catholic (r = 0.46):** Cantons with a larger Catholic population show moderately higher fertility, plausibly reflecting religious norms around family size.
- **Fertility ~ Infant.Mortality (r = 0.42):** Higher infant mortality goes hand in hand with higher fertility, as families may have more children when child survival is less certain.

### Agriculture

- **Agriculture ~ Examination (r = −0.69):** More agrarian cantons tend to score lower on the army examination, linking occupation structure to measured recruit quality.
- **Agriculture ~ Education (r = −0.64):** Cantons reliant on agriculture generally have lower literacy/education levels.
- **Agriculture ~ Catholic (r = 0.40):** Agricultural cantons are moderately more likely to be predominantly Catholic, reflecting the historical geography of Switzerland.
- **Agriculture ~ Infant.Mortality (r = −0.06):** There is essentially no linear relationship between the share of agricultural workers and infant mortality.

### Examination

- **Examination ~ Education (r = 0.70):** This is the strongest positive association in the table: cantons with better-educated populations also produce higher army-examination scores.
- **Examination ~ Catholic (r = −0.57):** Predominantly Catholic cantons tend to have lower examination scores, often attributed to language and schooling differences of the period.
- **Examination ~ Infant.Mortality (r = −0.11):** Better examination performance is only weakly (and not meaningfully) related to lower infant mortality.

### Education

- **Education ~ Catholic (r = −0.15):** The relationship between education and the share of Catholics is weakly negative and close to negligible.
- **Education ~ Infant.Mortality (r = −0.10):** Education and infant mortality show almost no linear association in these data.

### Catholic

- **Catholic ~ Infant.Mortality (r = 0.18):** Cantons with more Catholics have slightly higher infant mortality, but the association is weak.

### Infant.Mortality

- Infant mortality’s remaining pairwise relationships are all covered above; its only moderately strong link is with **Fertility (r = 0.42)**.
